### 2019-04-23 Sections/economics_of_the_quantum_internet.tex
ZKP++

\famousquote{If you put the federal government in charge of the Sahara Desert, in 5 years there'd be a shortage of sand.}{Milton Friedman}
\newline

@@

\sectionby{Scott Harrison}

@@

\sectionby{Scott Harrison \& Peter Rohde}

>>>>

\famousquote{If you put the federal government in charge of the Sahara Desert, in five years there'd be a shortage of sand.}{Milton Friedman}
\newline

@@

\sectionby{Scott Harrison}\index{Scott Harrison}

@@

\sectionby{Scott Harrison \& Peter Rohde}\index{Scott Harrison}

### 2019-04-23 Sections/introduction.tex
ZKP++

Perhaps the most exciting prospect in the quantum age is the development of quantum computers\index{Quantum computing}. Richard Feynman \cite{bib:Feynman85} was the first to ask the question \textit{``If quantum systems are so exponentially complex that we are unable to simulate them on our classical computers, can those same quantum systems be exploited in a controlled way to exponentially outperform our classical computers?"}\index{Richard Feynman}. Subsequently, the Deutsch-Jozsa algorithm \cite{bib:DeutschJozsa92}\index{Deutsch-Jozsa algorithm} demonstrated for the first time that algorithms can run on a quantum computer, exponentially outperforming any classical algorithm. Since then, an enormous amount of research has been dedicated to finding new quantum algorithms, and the search has indeed been a very fruitful one\footnote{See the Quantum Algorithm Zoo\index{Quantum Algorithm Zoo} for a comprehensive summary of the current state of knowledge on quantum algorithms (\url{http://math.nist.gov/quantum/zoo/}).}, with many important applications having been found, including, amongst many others\index{Quantum algorithms}:

>>>>

Perhaps the most exciting prospect in the quantum age is the development of quantum computers\index{Quantum computing}. Richard Feynman \cite{bib:Feynman85} was the first to ask the question \textit{``If quantum systems are so exponentially complex that we are unable to simulate them on our classical computers, can those same quantum systems be exploited in a controlled way to exponentially outperform our classical computers?''}\index{Richard Feynman}. Subsequently, the Deutsch-Jozsa algorithm \cite{bib:DeutschJozsa92}\index{Deutsch-Jozsa algorithm} demonstrated for the first time that algorithms can run on a quantum computer, exponentially outperforming any classical algorithm. Since then, an enormous amount of research has been dedicated to finding new quantum algorithms, and the search has indeed been a very fruitful one\footnote{See the Quantum Algorithm Zoo\index{Quantum Algorithm Zoo} for a comprehensive summary of the current state of knowledge on quantum algorithms (\url{http://math.nist.gov/quantum/zoo/}).}, with many important applications having been found, including, amongst many others\index{Quantum algorithms}:

### 2019-04-23 Sections/verification_of_cloud_quantum_computing.tex
ZKP++

Works for all \textbf{NP}-complete problems.

Alg.~\ref{alg:ZKP_graph}

Fig.~\ref{fig:ZKP_graph}

\begin{figure*}[!htpb]
	\includegraphics[clip=true, width=\textwidth]{ZKP_graph_isomorphism}
\captionspacefig \caption{The isomorphisms \mbox{$G_1\sim G_2$} (top), and \mbox{$G_1\sim H$} and \mbox{$G_2\sim H$} (bottom). Knowing the latter two isomorphisms simultaneously implies knowledge of \mbox{$G_1\sim G_2$} via composition of the permutations. However, knowing only one of them does not, since $H$ is chosen randomly. By repeatedly proving knowledge of one of the isomorphisms with $H$ we achieve asymptotic certainty that the prover must have known \mbox{$G_1\sim G_2$}, without actually revealing the associated permutation.}\label{fig:ZKP_graph}
\end{figure*}

\begin{table}[!htbp]
\begin{mdframed}[innertopmargin=3pt, innerbottommargin=3pt, nobreak]
\texttt{
function ZKP.GraphIsomorphism($G_1$, $G_2$):
\begin{enumerate}
	\item Graphs $G_1$ and $G_2$ are known to both verifier Victor, and prover Peggy.
	\item Peggy knows the permutation $\pi$ for the isomorphism \mbox{$G_1\sim G_2$},
	\begin{align}
		G_1 = \pi \cdot G_2.
	\end{align}
	\item Peggy wishes to prove to Victor that she knows $\pi$, without disclosing what it is.
	\item Peggy chooses another random permutation $\tilde\pi$, and constructs the new permuted graph $H$,
	\begin{align}
		H &= {\tilde\pi}\cdot G_1,\nonumber\\
		H &= {\tilde\pi}\cdot \pi \cdot G_2.
	\end{align}
	\item Peggy shares $H$ with Victor, randomly isomorphic to both $G_1$ and $G_2$.
	\item Victor randomly (\mbox{$p=1/2$}) asks Peggy to prove \textit{either} \mbox{$H\sim G_1$} or \mbox{$H\sim G_2$}.
	\item She reveals either $\tilde\pi$ or \mbox{$\tilde\pi\cdot\pi$} to Victor. He can now efficiently verify either \mbox{$H\sim G_1$} or \mbox{$H\sim G_2$} respectively, by performing the inverse permutation,
	\begin{align}
		G_1 &= {\tilde\pi}^{-1} \cdot H,\nonumber\\
		G_2 &= ({\tilde\pi}\cdot \pi)^{-1} \cdot H.
	\end{align}
	\item Victor is unable to determine $\pi$ from either scenario alone, but could if knew both isomorphisms simultaneously, since,
	\begin{align}
		\pi = {\tilde\pi}^{-1}\cdot (\tilde\pi\cdot\pi).
	\end{align}
	\item The above is repeated $n$ times. Each time, Peggy chooses a new random $\tilde\pi$.
	\item If Peggy does not actually know $\pi$, the probability of fraudulently passing this test $n$ times is,
	\begin{align}
		P_\text{deceive} = \frac{1}{2^n}	.
	\end{align}
	\item With confidence \mbox{$1-P_\text{deceive}$}, Victor knows that Peggy knows $\pi$ for the isomorphism \mbox{$G_1\sim G_2$}, without ever knowing $\pi$ itself.
	\item $\Box$
\end{enumerate}}
\end{mdframed}
\captionspacealg \caption{A zero-knowledge proof for the graph isomorphism problem, an \textbf{NP}-complete problem. Victor (verifier) provides two graphs, $G_1$ and $G_2$, to Peggy (prover), who can demonstrate with asymptotic certainty that she knows the isomorphism \mbox{$G_1\sim G_2$}, without disclosing the permutation $\pi$ that relates them.} \label{alg:ZKP_graph}\index{Zero-knowledge proofs}\index{Graph isomorphism}
\end{table}

>>>>

\comment{Ryan's links for NEXP ZKPs.}

\comment{General results.}

\comment{Ryan review.}

A \textit{zero-knowledge proof} (ZKP) is an interactive protocol\index{Interactive protocols} between two parties -- a \textit{prover}\index{Prover} Peggy, and a \textit{verifier}\index{Verifier} Victor -- where Peggy wishes to efficiently prove to Victor that she knows the solution to a problem, without actually revealing it. Thus, a ZKP can serve as a signature that a problem has been faithfully solved, without disclosing the solution.

ZKPs are useful in a number of cryptographic applications, most notably in authentication\index{Authentication}. In the case of classical computing, a variety of free software packages are available for compiling ZKPs for generic code. However, ZKPs become far more valuable in the case of cloud quantum computing, where there is an inherent complexity asymmetry between the client Victor (classical resources only) and the server Peggy (full quantum resources), whereby efficient ZKPs become a useful transactional tool during computational outsourcing. In a commercial context, a client can convince themselves that a server has actually performed an outsourced computation before finalising the transaction.

A conceptually simple example for illustrating the operation of a ZKP protocol is the \textit{graph isomorphism problem}\index{Graph isomorphism}. Two graphs are \textit{isomorphic}, denoted \mbox{$G_1\sim G_2$}, if there exists a permutation\index{Permutation} \mbox{$\pi\in S_n$}\footnote{In group theory\index{Group theory}, $S_n$ denotes the symmetric group\index{Symmetric group}, the set of all permutations on $n$ elements, of which there are \mbox{$|S_n|=n!$} (the order of the group).} on their vertex labels that makes them equivalent, i.e \mbox{$G_1=\pi\cdot G_2$}\footnote{Here we have employed the operator notation that \mbox{$\pi\cdot G$} means `permutation $\pi$ applied to graph $G$'. Alternately, in matrix notation, where $\pi$ is a permutation matrix and $G$ is an adjacency matrix\index{Adjacency matrix}, this operation implies the matrix conjugation\index{Conjugation} \mbox{$\pi\cdot G\cdot\pi^\top$}.}. The graph isomorphism problem is to find $\pi$ for arbitrary $G_1$ and $G_2$.

This problem is clearly contained in \textbf{NP}, since permuting vertices in graphs and directly comparing them are both computationally straightforward, making verification of the problem a polynomial-time affair. However, it is believed that explicitly determining the respective permutation is difficult in general. This is intuitively unsurprising, since for a graph with $n$ vertices, there are $n!$ possible permutations to consider (which is super-exponential), and therefore a na{\" i}ve brute-force\index{Brute-force} approach would require $O(n!)$ comparisons in the worst-case\footnote{This is the worst-case scenario. In many special cases, knowledge of underlying graph structure can simplify this enormously, yielding classically efficient runtimes.}. This problem is not believed to be contained in either \textbf{P} or \textbf{NP}-complete, and therefore presumed to be \textbf{NP}-intermediate\index{\textbf{NP}-intermediate} (Sec.~\ref{sec:introduction}).

\if 2\pubmode
\begin{figure}[!htpb]
	\includegraphics[clip=true, width=0.475\textwidth]{ZKP_graph_isomorphism}
\captionspacefig \caption{The isomorphisms \mbox{$G_1\sim G_2$} (top), and \mbox{$G_1\sim H\sim G_2$} (bottom). Coloured lines indicate the vertex relabelings associated with the respective isomorphisms. Knowing the latter two isomorphisms simultaneously implies knowledge of \mbox{$G_1\sim G_2$} via composition of the permutations. However, knowing only one of them does not, since $H$ is chosen randomly. By repeatedly proving knowledge of one of the isomorphisms with $H$ we achieve asymptotic certainty that the prover must have known \mbox{$G_1\sim G_2$}, without actually revealing the associated permutation.}\label{fig:ZKP_graph}
\end{figure}
\else
\begin{figure*}[!htpb]
	\includegraphics[clip=true, width=0.8\textwidth]{ZKP_graph_isomorphism}
\captionspacefig \caption{The isomorphisms \mbox{$G_1\sim G_2$} (top), and \mbox{$G_1\sim H\sim G_2$} (bottom). Coloured lines indicate the vertex relabelings associated with the respective isomorphisms. Knowing the latter two isomorphisms simultaneously implies knowledge of \mbox{$G_1\sim G_2$} via composition of the permutations. However, knowing only one of them does not, since $H$ is chosen randomly. By repeatedly proving knowledge of one of the isomorphisms with $H$ we achieve asymptotic certainty that the prover must have known \mbox{$G_1\sim G_2$}, without actually revealing the associated permutation.}\label{fig:ZKP_graph}
\end{figure*}
\fi

In the example isomorphism presented in Fig.~\ref{fig:ZKP_graph}(top), the vertex permutation mapping $G_1$ to $G_2$ could be expressed in vector form as,
\begin{align}
\pi = \begin{pmatrix}
	1 \\
	2 \\
	3 \\
	4 \\
	5
\end{pmatrix}_{G_1} \to \begin{pmatrix}
	1 \\
	2 \\
	4 \\
	3 \\
	5
\end{pmatrix}_{G_2},
\end{align}
where indices denote vertex labels, or equivalently via the permutation matrix,
\begin{align}
\pi = \begin{pmatrix}
	1 & 0 & 0 & 0 & 0 \\
	0 & 1 & 0 & 0 & 0 \\
	0 & 0 & 0 & 1 & 0 \\
	0 & 0 & 1 & 0 & 0 \\
	0 & 0 & 0 & 0 & 1
\end{pmatrix}.
\end{align}

\begin{table}[!htbp]
\begin{mdframed}[innertopmargin=3pt, innerbottommargin=3pt, nobreak]
\texttt{
function ZKP.GraphIsomorphism($G_1$, $G_2$):
\begin{enumerate}
	\item Graphs $G_1$ and $G_2$ are known to both verifier Victor, and prover Peggy.
	\item Peggy knows the permutation $\pi$ for the isomorphism \mbox{$G_1\sim G_2$},
	\begin{align}
		G_1 = \pi \cdot G_2.
	\end{align}
	\item Peggy wishes to prove to Victor that she knows $\pi$, without disclosing what it is.
	\item Peggy chooses another random permutation $\tilde\pi$, and constructs the new permuted graph $H$,
	\begin{align}
		H &= {\tilde\pi}\cdot G_1,\nonumber\\
		H &= {\tilde\pi}\cdot \pi \cdot G_2.
	\end{align}
	\item Peggy shares $H$ with Victor, randomly isomorphic to both $G_1$ and $G_2$.
	\item Victor randomly (\mbox{$p=1/2$}) asks Peggy to prove \textit{either} \mbox{$H\sim G_1$} or \mbox{$H\sim G_2$}.
	\item She accordingly reveals either $\tilde\pi$ or \mbox{$\tilde\pi\cdot\pi$} to Victor. He can now efficiently verify either \mbox{$H\sim G_1$} or \mbox{$H\sim G_2$} respectively, by performing the inverse permutation,
	\begin{align}
		G_1 &= {\tilde\pi}^{-1} \cdot H,\nonumber\\
		G_2 &= ({\tilde\pi}\cdot \pi)^{-1} \cdot H.
	\end{align}
	\item Victor is unable to determine $\pi$ from either scenario alone, but could were he to know \textit{both} isomorphisms simultaneously, since,
	\begin{align}
		\pi = {\tilde\pi}^{-1}\cdot (\tilde\pi\cdot\pi).
	\end{align}
	\item The above is repeated $n$ times. Each time, Peggy chooses a new random $\tilde\pi$.
	\item If Peggy does not actually know $\pi$, the probability of fraudulently passing this test $n$ times is,
	\begin{align}
		P_\text{deceive} = \frac{1}{2^n}	.
	\end{align}
	\item With confidence \mbox{$1-P_\text{deceive}$}, Victor knows that Peggy knows $\pi$, without knowing it himself.
	\item $\Box$
\end{enumerate}}
\end{mdframed}
\captionspacealg \caption{A zero-knowledge proof for the graph isomorphism problem. Victor (verifier) provides two graphs to Peggy (prover), who can demonstrate with asymptotic certainty that she knows their isomorphism, without disclosing the associated permutation relating them.} \label{alg:ZKP_graph}\index{Zero-knowledge proofs}\index{Graph isomorphism}
\end{table}

To provide a ZKP for this, Peggy's goal is to prove that she knows $\pi$, without explicitly revealing it. An efficient, randomised, classical ZKP protocol for achieving this is provided in Alg.~\ref{alg:ZKP_graph}, and a specific example illustrated in Fig.~\ref{fig:ZKP_graph}. The key underlying principle here is to obscure $\pi$ through randomisation, whereby instead of directly proving knowledge of $\pi$, it is implied via multiple proofs of isomorphisms with intermediate random graphs.
